### 2022-01-26 paper/paper.tex
methodology: expound on algorithm, add backend and testing

@@

The experimental setup and configuration will be composed of three elements: the
camera, the keyboard, and the environment.

\subsection{Camera}
The camera setup will use a single monocular camera positioned in a top-down
view. The camera should capture the entirety of the keyboard and the movement of
the ten fingers. Figure~\ref{fig:metho-keeb} shows what the camera will capture
once placed in its correct position.

The specific camera to be used will be a Logitech C270.
Figure~\ref{fig:metho-cam} is a picture of this camera. This is a 0.9 mega pixel
webcam with a CMOS sensor. This camera is capable of capturing color videos in
720p/30fps and it has a diagonal field of view of 55\degree. This camera has a
universal mounting clip that will reduce the difficulty in mounting the camera
\parencite{logitech}.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\textwidth]{webcam.png}
	\caption{Logitech C270. Reprinted from \fullcite{logitech}}
	\label{fig:metho-cam}
	\centering
\end{figure}

\subsection{Keyboard}
\label{section:metho-keeb}
The keyboard will be a 60\% keyboard as shown in Figure~\ref{fig:metho-keeb}.
This will limit the necessary mapping for the algorithm to the alphanumeric
portion of the keyboard. This keyboard choice will also lessen the area that the
camera will need to capture, as this keyboard type is considerably smaller
compared to a full-size keyboard. The keyboard will have key caps and a case in
a light color that will contrast the dark surface that it will be placed on.
This is to improve initial keyboard detection. In addition, the keyboard layout
will also be ANSI, due to the availability and widespread adoption of the layout
in the Philippines.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
	\caption{Keyboard, camera angle, and placement to be used for testing.}
	\label{fig:metho-keeb}
\end{figure}

@@

\subsubsection{Convert to Grayscale}
The frame will then be converted to a 256 level grayscale image. This step will
be performed because future steps of the algorithm will not require color values
to work. In addition, the performance of these steps will be improved.

\subsubsection{Edge Detection using Sobel Filter}
A Sobel Filter is an edge detection filter that uses a 3$\times$3 kernel that is
convolved twice. Once horizontally, and another vertically to produce a
grayscale image of the outlines within the frame. The kernels used by the Sobel
Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.
This filter will be used in the algorithm to obtain the four edges of the
keyboard.

@@

\subsubsection{Find Contours (Simple)}
\label{section:metho-algo-key-contours}
Contours are curves joining all continuous points that have the same color or
intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
to be used to find the contours of the outlines of the object found using the
previous step. This OpenCV function implements the algorithm of \cite{contours}.
The expected result is four contour points which are positioned on the four
edges of the keyboard.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{image-map-final.png}
	\caption{Expected output of the algorithm}
	\label{fig:metho-algo-key-overlay}
	\centering
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{image-map-half.png}
	\caption{Overlaid image map shown at 60\% opacity}
	\label{fig:metho-algo-key-half}
	\centering
\end{figure}

\subsection{Computer Vision based Finger Detection, and Tracking}
\label{section:metho-algo-finger}

A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video. The specific algorithm used for finger
detection and tracking will be the solution offered by MediaPipe dubbed
MediaPipe Hands. This algorithm is composed of two ML models working in
conjunction to be able to detect the different parts of the hands and track them
accurately.

@@

MediaPipe Hands detects the palms first, instead of the whole hands with one
model because the hands lack high contrast patterns. This prevents the model to
detect the hand accurately. In addition, detecting a palm is simpler compared to
detecting hands with articulated fingers since estimating a bounding box around
a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
modeled using only square anchors reducing the number of anchors by a factor of
3--5 \parencite{mediapipe-hands}.

@@

\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
Since there is a known color-key mapping based on the Virtual Keyboard Map, the
algorithm will find the correct color to search for based on the received
keypress packet. This color is isolated from the Virtual Keyboard Map, and the
region of interest and its coordinates are obtained. The edges of this ROI will
then be obtained by finding its contours. The specific implementation if this is
explained in Section~\ref{section:metho-algo-key-contours}.

@@

\subsubsection{Find which fingertip is on that ROI}
The Finger Tracking Frame will contain the pixel positions of each landmark of
the hand. For this step, the algorithm will find the landmark which is
positioned within the region of interest obtained from the previous step. This
will be done using a series of checks:

\subsubsection{Return which fingertip}
The landmark will then be used to determine which specific finger corresponds to
the keypress. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
possible landmark that may be returned from the previous step. This step will
then simply return the name of the landmark, up until the first underscore. As
an example, if the landmark found is \emph{MIDDLE\_FINGER\_DIP}, this step will
return \emph{MIDDLE} denoting that the middle finger is the finger that
corresponds with the keypress.

@@

@@

\subsection{UX/UI design and Front-end development for the touch typing trainer}
The touch typing trainer will have its UX/UI design created in Figma. This
design will be based on other type tests such as Monkeytype
\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
for showing real-time finger-key identification and mapping.
Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
the design. Additional screens, such as historical statistics, in-platform
tutorial, and a written guide are shown in the appendix.

@@

@@

The front-end will be web based, and the framework the front-end will be
developed in will be Svelte, due to its community resources, performance, and
the researcher's familiarity with the framework.

\subsection{Back-end development of a touch typing trainer using the program created
	in \ref{section:metho-algo} and the experimental setup identified in
	\ref{section:metho-setup}}
\subsection{Testing of finger-key identification and mapping accuracy}
\section{Metric}
%% 4. Finger Placement Keyboard Typing Metric Development
%%     1. Determine if the metric is for individual key presses or for the whole typing sequence or both.
%%     2. Evaluate which variables need to be taken into account for the metric depending on (5.a):
%%         1. Actual Finger Used
%%         2. Correct Finger to be Used
%%         3. Specific key pressed
%%         4. Number of keys pressed with the correct fingers
%%         5. Number of keys pressed with the wrong fingers

>>>>

% Adjust Figure Spacing
\setlength\intextsep{1cm}

@@

The experimental setup and configuration will be composed of three elements: the
camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup-cam}
shows the sketch of the complete experimental setup.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\textwidth]{webcam.png}
	\caption{Sketch of the experimental setup}
	\label{fig:metho-setup-cam}
	\centering
\end{figure}

\subsection{Camera}
The camera setup will use a single monocular camera positioned in a top-down
view. The camera will capture the entirety of the keyboard and the movement of
the ten fingers. To do so, it will be mounted on top of the monitor and the
camera will point down towards the table. Figure~\ref{fig:metho-setup-keeb}
illustrates what the camera will capture once placed in its correct position.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
	\caption{Keyboard, camera angle, and placement to be used for testing.}
	\label{fig:metho-setup-keeb}
\end{figure}

The specific camera to be used will be a Logitech C920.
Figure~\ref{fig:metho-setup-cam} is a picture of this camera. It is a 3 mega pixel
webcam that is capable of capturing color video in 1080p/30fps and 720p/30fps
with a diagonal field of view of 78\degree. This camera has a universal mounting
clip that will allow the camera to be correctly positioned within the
experimental setup \parencite{logitech}.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\textwidth]{webcam.png}
	\caption{Logitech C920. Reprinted from \fullcite{logitech}}
	\label{fig:metho-setup-cam}
	\centering
\end{figure}

\subsection{Keyboard}
\label{section:metho-keeb}

The keyboard will be a 60\% keyboard as shown in
Figure~\ref{fig:metho-setup-keeb}. This will limit the necessary mapping for the
algorithm to the alphanumeric portion of the keyboard. This keyboard choice will
also lessen the area that the camera will need to capture, as this keyboard type
is considerably smaller compared to a full-size keyboard. The keyboard will have
key caps and a case in a light color that will contrast the dark surface that it
will be placed on. This is to improve initial keyboard detection. In addition,
the keyboard layout will also be ANSI, due to the availability and widespread
adoption of the layout in the Philippines.

@@

\subsubsection{Convert to Grayscale}
The frame will be converted to a 256 level grayscale image. This step will be
performed because future steps of the algorithm will not require color values to
work. In addition, the performance of the following steps will be also improved
as the number of dimensions to be analyzed is reduced.

\subsubsection{Edge Detection using Sobel Filter}
A Sobel Filter is an edge detection filter that uses a 3$\times$3 kernel that is
convolved twice. Once horizontally, and another vertically to produce a
grayscale image of the outlines within the frame. The kernels used by the Sobel
Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.

@@

The expected result of this step is a black and white image that highlights the
edges of the keyboard in white.

\subsubsection{Find Contours (Simple)}
\label{section:metho-algo-key-contours}
Contours are curves joining all continuous points that have the same color or
intensity \parencite{opencv-contours}. In this algorithm the OpenCV function
\texttt{findContours} will be used to find the contours of the outlines of the
object found using the previous step. This function accepts a contour
approximation method as one of its arguments. Two methods are provided by
OpenCV, \texttt{CHAIN\_APPROX\_NONE} and \texttt{CHAIN\_APPORX\_SIMPLE}. The
former returns all contours in the shape, while the latter removes redundant
points and returns the least amount of points that describes the shape. The
algorithm will use the latter, as only the extreme edges of the keyboard needs
to be detected. This OpenCV function implements the algorithm of
\cite{contours} in their paper ``\citefield{contours}{title}''.

The expected result of this algorithm is four contour points which are
positioned on the four edges of the keyboard.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{image-map-final.png}
	\caption{Expected output of the algorithm}
	\label{fig:metho-algo-key-overlay}
	\centering
\end{figure}

\hfill

\begin{figure}[H]
	\centering
	\includegraphics[width=1\textwidth]{image-map-half.png}
	\caption{Overlaid image map shown at 60\% opacity}
	\label{fig:metho-algo-key-half}
	\centering
\end{figure}

\subsection{Computer Vision based Finger Detection, and Tracking}
\label{section:metho-algo-finger}
A computer vision algorithm will be chosen as a starting point for detecting,
and tracking fingers within a video. The specific algorithm used for finger
detection and tracking will be the solution offered by MediaPipe dubbed
MediaPipe Hands. This algorithm is composed of two ML models working in
conjunction to be able to detect the different parts of the hands and track them
accurately.

@@

MediaPipe Hands detects the palms first, instead of the whole hands with one
model because the hands lack high contrast patterns. This reduces the model's
ability to detect the hand with accuracy. In addition, detecting a palm is
simpler compared to detecting hands with articulated fingers since estimating a
bounding box around a rigid objects, i.e.\ a palm, is much simpler. Furthermore,
a palm can be modeled using only square anchors reducing the number of anchors
by a factor of 3--5 \parencite{mediapipe-hands}.

@@

\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
Since there is a known color-key mapping based on the Virtual Keyboard Map, the
algorithm will find the correct color to search for based on the received
keypress packet. This color is isolated from the Virtual Keyboard Map, and the
region of interest and its coordinates are obtained. The edges of this ROI will
then be obtained by finding its contours. The specific implementation of this is
explained in Section~\ref{section:metho-algo-key-contours}.

@@

\subsubsection{Find which fingertip is on that ROI}
The Finger Tracking Frame will contain the pixel positions of each landmark of
the hand. For this step, the algorithm will find the landmark which is
positioned within the region of interest obtained from the previous step. This
will be done using a series of checks.

Each landmark's coordinates will be compared to the coordinates of the edges of
the ROI. A landmark will be determined as positioned within the ROI if all of the
following conditions are true:
\begin{enumerate}
	\item In the X axis, the landmark's coordinates is greater than one or both of
		the two coordinates found of the left side of the ROI
	\item In the X axis, the landmark's coordinates is less than one or both of
		the two coordinates found of the right side of the ROI
	\item In the Y axis, the landmark's coordinates is greater than one or both of
		the two coordinates found of the top side of the ROI
	\item In the Y axis, the landmark's coordinates is less than one or both of
		the two coordinates found of the bottom side of the ROI
\end{enumerate}

These conditions maximize the total area of the ROI and is not strict about
exact accuracy. In essence, these conditions will create a perfectly rectangular
box that contains the quadrilateral formed by the four points from the previous
step.

\subsubsection{Return which fingertip}
The landmark will then be used to determine which specific finger corresponds to
the keypress. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
possible landmark that may be returned from the previous step. This step will
then return the name of the landmark, up until the first underscore. As an
example, if the landmark found is \texttt{MIDDLE\_FINGER\_TIP}, this step will
return \texttt{MIDDLE} denoting that the middle finger is the finger that
corresponds with the keypress. In addition, the specific hand will also be
returned as one of two strings, \texttt{Left} and \texttt{Right}, since
this information is also bundled together with the landmarks.

@@

\subsection{Implementation}
\label{section:metho-algo-implementation}
The previously discussed algorithms will be implemented using Python. The
Keyboard Detection, and Mapping Algorithm in
Section~\ref{section:metho-algo-keyboard} will be implemented as a separate
submodule in Python using OpenCV.\@The Finger Detection, and Tracking Solution
in Section~\ref{section:metho-algo-finger} will be consumed using the prebuilt
Python package offered by the MediaPipe team. Finally, the integration of the
two will be done as another separate submodule in Python.

@@

\subsection{UX/UI design and Front-end development for the touch typing trainer}
\subsubsection{Design}
This design will be based on other type tests such as Monkeytype
\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
for showing real-time finger-key identification and mapping.
Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
the design. Additional screens, such as historical statistics, in-platform
tutorial, and a written guide are shown in the appendix.

@@

Figure~\ref{fig:ui-test} has four main components. From the top, the first
component is the navigation bar. This directs the user to the different pages
within the website. The second component in the upper left displays the number
of remaining words to be typed. This is 16 in the example. Below that is the
test sequence. This shows the user what characters to type, and highlights
mistakes in two distinct colors. Red if the character pressed was incorrect and
purple if the character typed was correct, but the finger used to type it is
wrong. The last main component in this screen is the keyboard that illustrates
the correct finger positioning, with each finger corresponding to a different
color.

@@

Figure~\ref{fig:ui-stat} has two main components. The first component at the top
shows the three major statistics for the typing test: Words per minute,
Accuracy, and Finger placement accuracy. Finger placement accuracy will be
explained in Section~\ref{label:metric}. The component below the statistics is a
graph plotting the major statistics of the user in time. This includes when the
user made the mistake and when the user slowed down, or increased their speed.

\subsubsection{Implementation}
The design will be created in Figma and the front-end will be web based.The
framework the front-end will be developed in will be Svelte, due to its
community resources, performance, and the researcher's familiarity with the
framework. The front-end will serve as a display for the information and the
interface which the user interacts and inputs information into the trainer. No
calculation will be done within the front-end.

\subsection{Back-end development of a touch typing trainer using the program created
	in \ref{section:metho-algo} and the experimental setup identified in
	\ref{section:metho-setup}}

The back-end will handle all of the computation for the trainer---mainly
finger-key identification. The backend will also handle authentication and data
storage.

The back-end will expose a REST API. This is a type of application programming
interface (API) that conforms to the REST architectural style. One key
characteristic of the this style is its statelessness and cacheable data
\parencite{rest}. This will be implemented using Django, a Python REST API
framework.

\subsubsection{Finger-key Identification}
The submodule that integrates all of the sub algorithms will be called within
the framework. The framework will then respond to requests for finger-key
identification using this submodule. This will allow the back-end to expose its
finger-key identification abilities and allow the front-end to use this data for
the trainer.

\subsubsection{Authentication}
The core authentication system will rely on JSON Web Tokens (JWTs) . According
to \cite{jwt}, ``JSON Web Token (JWT) is a compact, URL-safe means of
representing claims to be transferred between two parties.'' This allows for a
user to claim that they are that specific user securely as the claim is
cryptographically signed. For the backend of the trainer, the JWT will be signed
using a secret and it will be encoded in HMAC SHA256.

\subsubsection{Data Storage}
The trainer will store all of the data in a relational database using
PostgreSQL.\@This specific technology was chosen due to its maturity, community,
and industry recommendations and the researchers' familiarity with the
technology.

\subsection{Testing of finger-key identification and mapping accuracy}

There will be 15 videos of the researcher performing the typing test sequences.
The researcher will then manually perform finger-key identification for all key
presses present in the video. The finger-key identification submodule will then
perform the same process for the same videos. The accuracy of the submodule will
then be obtained by identifying which of its identification is incorrect, based
on the manual identification.

\section{Metric}
\label{section:metric}

There will be two total metrics obtained pertaining to finger and key mapping.
The first will be per test sequence, and the second will be per key.

%% 4. Finger Placement Keyboard Typing Metric Development
%%     1. Determine if the metric is for individual key presses or for the whole typing sequence or both.
%%     2. Evaluate which variables need to be taken into account for the metric depending on (5.a):
%%         1. Actual Finger Used
%%         2. Correct Finger to be Used
%%         3. Specific key pressed
%%         4. Number of keys pressed with the correct fingers
%%         5. Number of keys pressed with the wrong fingers
